Exploratory: Optimal Resolution

### 2021-05-08 exploratory/exploratory.Rmd
ms updates

---
title: "Exploratory: Optimal Resolution"
author: "Courtney R Armour"
output: html_document
---
updated: `r Sys.Date()`
```{r setup_environment, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE)
options(scipen=999)

@@

## Table of Contents
- [Input data overview](#input-data-overview)
- [Hyperparameter Performance](#hyperparameter-performance)
- [Model Performance](#model-performance)
  - [Random Forest](#random-forest-model-performance-with-significance)
- [DADA2 Comparison](#dada2-comparison)
- [Feature Importance](#feature-importance)

## Input data overview

To total number of samples and features before and after preprocessing for each taxnomic level. 
```{r input}

@@

## Hyperparameter Performance

@@

### Random Forest
The default hyperparameter (mtry) selection for RF is:

1. sqrt_features / 2
1. sqrt_features
1. sqrt_features * 2

@@

Two metrics: alpha and lambda  
By default alpha is set to zero (for L2 regularization) and lambda values are: 1e-04, 1e-03, 1e-02, 1e-01, 1e+00, 1e+01
I added 1e-05 as a lambda value for phylum and class, and higher values of lambda for the rest

@@

## Model Performance

@@

### DADA2 Comparison

@@

### Random Forest Model Performance with Significance

@@

<div class = "row">
<div class = "col-md-6">
```{r rf_table,echo=FALSE}
t_table %>%
  select(level,rf) %>% 
  mutate(rf=round(rf,digits = 3)) %>% 
  rename("Median AUC"=rf,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```

</div>

<div class = "col-md-6">
```{r rf_plot,echo=F,message=F}
m <- "Random Forest"

@@

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method)) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.2,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  #stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")
ggsave("./figures/rf_with_stats.png",width = 5,height=4,units="in")

```
</div>
</div>

```{r rf_wstat_side,include=F}
sub_stats <- sub_stats %>% 
  #mutate(y.position=seq.int(0.85,.95,length.out = nrow(sub_stats)))
  mutate(y.position=c(0.85,0.95,0.9,.95))

ggplot(sub_df,aes(x=factor(level,levels=c("ASV", "OTU","Genus","Family","Order","Class", "Phylum")),
                           y=AUC)) + 
  #geom_jitter(aes(color=method,width=0,height=0.2)) +
  geom_boxplot(aes(fill=method,alpha=0.8)) +
  scale_fill_manual(values = pal2[m]) +
  #scale_color_manual(values = pal2[m]) +
  #scale_x_continuous(limits=c(0.4,1)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  ggtitle(m) +
  stat_pvalue_manual(sub_stats,label="sig",coord.flip = T,angle = 0,hjust=0) +
  coord_flip() +
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") 
ggsave("./figures/rf_with_stats_sideways.png",width = 5,height=4,units="in")
```

@@

<div class = "row">
<div class = "col-md-6">
```{r glmnet_table,echo=FALSE}
t_table %>%
  select(level,glmnet) %>% 
  mutate(glmnet=round(glmnet,digits = 3)) %>% 
  rename("Median AUC"=glmnet,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
</div>

<div class = "col-md-6">
```{r glmnet_plot,echo=F,message=F}
## glmnet Model Performance
m <- "Logistic Regression"

@@

```
</div>
</div>

@@

<div class = "row">
<div class = "col-md-6">
```{r rpart2_table,echo=FALSE}
t_table %>%
  select(level,rpart2) %>% 
  mutate(rpart2=round(rpart2,digits = 3)) %>% 
  rename("Median AUC"=rpart2,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
</div>

<div class = "col-md-6">
```{r rpart2_plot,echo=F,message=F}
## rpart2 Model Performance
m <- "Decision Tree"

@@

<div class = "row">
<div class = "col-md-6">
```{r svmRadial_table,echo=FALSE}
t_table %>%
  select(level,svmRadial) %>% 
  mutate(svmRadial=round(svmRadial,digits = 3)) %>% 
  rename("Median AUC"=svmRadial,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
</div>

<div class = "col-md-6">
```{r svmRadial_plot,echo=F,message=F}
## svmRadial Model Performance
m <- "SVM Radial"

@@

<div class = "row">
<div class = "col-md-6">
```{r xgbTree_table,echo=FALSE}
t_table %>%
  select(level,xgbTree) %>% 
  mutate(xgbTree=round(xgbTree,digits = 3)) %>% 
  rename("Median AUC"=xgbTree,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
</div>

<div class = "col-md-6">
```{r xgbTree_plot,echo=F,message=F}
## XGBoost Model Performance
m <- "XGBoost"

@@

```
</div>
</div>

## Feature Importance
The mikropml packages includes an option for finding [feature importance](http://www.schlosslab.org/mikropml/reference/get_feature_importance.html) using a permutation method. This significantly increases the time to run the models. The function permutes each model feature (e.g. genus) and recalculates the AUC with that feature permuted. The outputs are the permuted AUC value and the difference between this AUC value and the actual AUC value. Larger positive differences can be interpreted as being more important since the AUC value is smaller when this feature is permuted.

@@

### Top 10 Important features
```{r importance_rank, message=F,echo=F}

@@

top_10_tax <- ranked_tax %>% 
  filter(rank <= 10) %>% 
  select(level,Taxonomy,rank) %>% 
  pivot_wider(names_from = level,values_from = Taxonomy)

kable(top_10_tax,format="html",booktabs=T,align="clllllll",
      col.names = c("Rank",levels_names)) %>% 
   kable_material(c("striped", "hover","condensed"))
```

## Prevalence

```{r prevalence_preprocess,echo=F,message=F}
get_prevalence_preprocess <- function(file,file_proc,level){
  data <- as.data.frame(fread(file))
  data_proc <- as.data.frame(fread(file_proc))
  keep_names <- str_replace(colnames(data_proc)[-1],"_1","")
  n_samp <- nrow(data_proc)
  prev <- select(data,keep_names) %>% 
    pivot_longer(everything(),names_to="name",values_to="count") %>% 
    filter(count > 0) %>% 
    group_by(name) %>% 
    tally() %>% 
    mutate(prevalence=n/n_samp) %>% 
    mutate(level := !!level) %>% 
    select(level,name,prevalence)
  return(prev)
}

# prev_df <- get_prevalence_preprocess("./data/final/phylum/input_data.csv",
#                                      "./data/final/phylum/input_data_preproc.csv",
#                                      "phylum") %>%
#   bind_rows(get_prevalence_preprocess("./data/final/class/input_data.csv",
#                                       "./data/final/class/input_data_preproc.csv",
#                                       "class")) %>%
#   bind_rows(get_prevalence_preprocess("./data/final/order/input_data.csv",
#                                       "./data/final/order/input_data_preproc.csv",
#                                       "order")) %>%
#   bind_rows(get_prevalence_preprocess("./data/final/family/input_data.csv",
#                                       "./data/final/family/input_data_preproc.csv",
#                                       "family")) %>%
#   bind_rows(get_prevalence_preprocess("./data/final/genus/input_data.csv",
#                                       "./data/final/genus/input_data_preproc.csv",
#                                       "genus")) %>%
#   bind_rows(get_prevalence_preprocess("./data/final/otu/input_data.csv",
#                                       "./data/final/otu/input_data_preproc.csv",
#                                       "otu")) %>%
#   bind_rows(get_prevalence_preprocess("./data/final/asv/input_data.csv",
#                                       "./data/final/asv/input_data_preproc.csv",
#                                       "asv"))
# write_csv(prev_df,"./data/analysis/exploratory3/prevalence_by_level_preprocess.csv")

prevalence_prepro <- read_csv("./data/analysis/exploratory3/prevalence_by_level_preprocess.csv")

prevalence_prepro$level <- factor(prevalence_prepro$level,levels=levels)

ggplot(prevalence_prepro,aes(x=level,y=prevalence,fill=level)) + 
  #geom_jitter() +
  geom_boxplot() +
  stat_summary(fun = mean, color="red",geom = "point", shape = 18, size = 4,show.legend = FALSE) +
  theme_bw() + xlab("") +
  theme(legend.position = "none",
        axis.text = element_text(size=14),
        axis.title.y = element_text(size=14))
ggsave("./data/analysis/exploratory3/prevalence_preprocess.png",width=5,height=4,units = "in")

ggplot(prevalence_prepro,aes(x=prevalence,fill=level)) +
  geom_histogram(binwidth = 0.05) +
  facet_wrap(.~level,scales = "free") +
  theme_bw() +
  scale_y_continuous(expand=c(0,0)) +
  scale_x_continuous(expand=c(0,0)) +
  theme(legend.position = "none") 
ggsave("./data/analysis/exploratory3/prevalence_histogram.png",width=5,height=4,units = "in")

ggplot(prevalence_prepro,)

tax_count <- prevalence_prepro %>% 
  group_by(level) %>% 
  mutate(numOTU = n_distinct(name)) %>% 
  select(level,numOTU) %>% 
  distinct()

# prevalence_prepro %>% 
#   group_by(level) %>% 
#   count(prevalence) %>% 
#   inner_join(tax_count,by="level") %>% 
#   mutate(pct_OTU = n/numOTU) %>% 
#   mutate(prevalence = round(prevalence,digits=2)) %>% 
#   ggplot(aes(x=prevalence,y=pct_OTU)) +
#     geom_line() +
#     facet_wrap(~level,scales = "free")
  #ggplot(aes(x=prevalence,y=n)) +
  #  geom_bar(stat="identity") +
  #  facet_grid(~level)

pct_bins <- NULL
level_nOTU <- NULL
for(l in levels){
  sub_data <- prevalence_prepro %>% 
    filter(level==l)
  hist <- hist(sub_data$prevalence,breaks = seq(0,1,0.25))

@@

level_nOTU <- level_nOTU %>% 
  mutate(level_name = levels_names) 
labs <- paste0(level_nOTU$level_name," (",level_nOTU$nOTU,")")
pct_bins %>%
  mutate(level=factor(level,levels=levels,labels=labs)) %>% 
  ggplot(aes(x=prev,y=pct,fill=level)) +
    geom_bar(stat = "identity") +
    facet_wrap(~level) +
    scale_y_continuous(expand=c(0,0),limits=c(0,0.8),labels = scales::percent_format()) +
    scale_x_continuous(expand=c(0,0),breaks = c(0,.25,0.50,.75,1)) +
    xlab("Prevalence in Samples") +
    ylab("Percent of Taxa") +
    theme_bw() +
    theme(legend.position = "none") +
  geom_text(aes(label = paste0((round(pct,digits=2)*100),"%")),nudge_y=0.05)
ggsave("./data/analysis/exploratory3/bin_prevalence.png")

# 
# ggplot(prevalence_prepro,aes(x=prevalence,fill=level)) +
#   geom_histogram(binwidth = 0.25) +
#   facet_wrap(.~level,scales = "free") +
#   theme_bw() +
#   scale_y_continuous(expand=c(0,0)) +
#   scale_x_continuous(expand=c(0,0)) +
#   theme(legend.position = "none")

>>>>

---
title: "Exploratory: Optimal Resolution"
author: "Courtney R Armour"
output: html_document
---

updated: `r Sys.Date()`

```{r setup_environment, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE)
options(scipen=999)

@@

## Table of Contents

-   [Input data overview](#input-data-overview)

-   [Hyperparameter Performance](#hyperparameter-performance)

-   [Model Performance](#model-performance)

    -   [Random Forest](#random-forest-model-performance-with-significance)

-   [DADA2 Comparison](#dada2-comparison)

-   [Feature Importance](#feature-importance)

-   [Prevalence](#prevalence)

## Input data overview {#input-data-overview}

To total number of samples and features before and after preprocessing for each taxnomic level.

```{r input}

@@

## Hyperparameter Performance {#hyperparameter-performance}

@@

### Random Forest

The default hyperparameter (mtry) selection for RF is:

1.  sqrt_features / 2
2.  sqrt_features
3.  sqrt_features \* 2

@@

Two metrics: alpha and lambda\
By default alpha is set to zero (for L2 regularization) and lambda values are: 1e-04, 1e-03, 1e-02, 1e-01, 1e+00, 1e+01 I added 1e-05 as a lambda value for phylum and class, and higher values of lambda for the rest

@@

## Model Performance {#model-performance}

@@

### DADA2 Comparison {#dada2-comparison}

@@

### Random Forest Model Performance with Significance {#random-forest-model-performance-with-significance}

@@

::: {.row}
::: {.col-md-6}
```{r rf_table,echo=FALSE}
t_table %>%
  select(level,rf) %>% 
  mutate(rf=round(rf,digits = 3)) %>% 
  rename("Median AUC"=rf,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
:::

::: {.col-md-6}
```{r rf_plot,echo=F,message=F}
m <- "Random Forest"

# sub_stats <- p.table.model %>% 
#    mutate(level1=factor(level1,levels=c("phylum","class","order","family","genus","otu","asv"),
#                             labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
#           level2=factor(level2,levels=c("phylum","class","order","family","genus","otu","asv"),
#                             labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
#          model=factor(model,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
#                        labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
#   filter(model==m) %>% 
#   select(-model) %>% 
#   pivot_wider(names_from=level2,values_from=p_value) %>% 
#   column_to_rownames(var="level1")
# sub_stats <- sub_stats[levels_names[1:6],levels_names[2:7]]

# sub_stats <- sub_stats %>% 
#   mutate(y.position=seq.int(0.85,1.05,length.out = nrow(sub_stats))) %>% 
#   rename(group1=level1,group2=level2)

ggplot(sub_df,aes(x=factor(level,levels=c("ASV", "OTU","Genus","Family","Order","Class", "Phylum")), y=AUC)) + 
#ggplot(sub_df,aes(x=level, y=AUC)) + 
  geom_jitter(aes(fill=method,color=method),height = 0, width = 0.2,shape=21) +
  stat_summary(fun.data = median_hilow,fun.args=0.5, geom="crossbar", width=0.5, color="black") +
  scale_fill_manual(values = alpha(pal2[m],0.6)) +
  scale_color_manual(values = pal2[m]) +
  scale_y_continuous(limits=c(0.45,0.87),breaks = seq(0.5,0.8,0.1),expand=c(0,0)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.y = element_blank(),
        axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  coord_flip() +
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  annotate("text",x=levels_names,y=rep(0.85,length(levels_names)),label=c("A","B","C","D,E","D","D","E"))
ggsave("./figures/rf_with_stats.png",width = 5,height=4,units="in")

```
:::
:::

```{r rf_wstat_options,include=F}

# OPTION 1 ------------
m <- "Random Forest"

@@

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  geom_boxplot(aes(fill=method,alpha=0.8)) +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.4,1.05)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  #ggtitle(m) +
  #stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")

# OPTION 2 ------------

ggplot(sub_df,aes(x=factor(level,levels=c("ASV", "OTU","Genus","Family","Order","Class", "Phylum")),
                           y=AUC)) + 
  #geom_jitter(aes(color=method,width=0,height=0.2)) +
  geom_boxplot(aes(fill=level,alpha=0.8)) +
  #scale_fill_manual(values = pal2[m]) +
  #scale_color_manual(values = pal2[m]) +
  #scale_x_continuous(limits=c(0.4,1)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  #ggtitle(m) +
  stat_pvalue_manual(sub_stats,label="sig",coord.flip = T,angle = 0,hjust=0) +
  coord_flip() +
  geom_hline(yintercept = 0.5,color="grey",lty="dashed")

# OPTION 3 ------------

sub_stats <- sub_stats %>% 
  #mutate(y.position=seq.int(0.85,1.05,length.out = nrow(sub_stats))) %>% 
  mutate(y.position=c(0.85,0.95,0.9,.95)) %>%
  rename(group1=level1,group2=level2)

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  #geom_boxplot(aes(fill=method)) +
  geom_jitter(aes(fill=method),height = 0, width = 0.2,alpha=0.8,shape=21) +
  stat_summary(fun.data = median_hilow,fun.args=0.5, geom="crossbar", width=0.5, color="black") +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.4,1),breaks = seq(0.4,1,0.1)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  #ggtitle(m) +
  #stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  stat_pvalue_manual(sub_stats,label="sig")

# OPTION 4 ------------

sub_stats <- sub_stats %>% 
  #mutate(y.position=seq.int(0.85,.95,length.out = nrow(sub_stats)))
  mutate(y.position=c(0.85,0.95,0.9,.95))

ggplot(sub_df,aes(x=factor(level,levels=c("ASV", "OTU","Genus","Family","Order","Class", "Phylum")),
                           y=AUC)) + 
  geom_jitter(aes(fill=level),height = 0, width = 0.2,alpha=0.8,shape=21) +
  stat_summary(fun.data = median_hilow,fun.args=0.5, geom="crossbar", width=0.5, color="black") +
  #scale_fill_manual(values = pal2[m]) +
  #scale_color_manual(values = pal2[m]) +
  #scale_x_continuous(limits=c(0.4,1)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  scale_y_continuous(position = "right") +
  #ggtitle(m) +
  stat_pvalue_manual(sub_stats,label="sig",coord.flip = T,angle = 0,hjust=0) +
  coord_flip() +
  geom_hline(yintercept = 0.5,color="grey",lty="dashed")

# OPTION 5 ------------

sub_stats <- p.table.model %>% 
   mutate(level1=factor(level1,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
          level2=factor(level2,levels=c("phylum","class","order","family","genus","otu","asv"),
                            labels=c("Phylum","Class","Order","Family","Genus","OTU","ASV")),
         model=factor(model,levels=c("rf","glmnet","xgbTree","svmRadial","rpart2"),
                       labels=c("Random Forest","Logistic Regression","XGBoost","SVM Radial","Decision Tree"))) %>% 
  filter(model==m) %>% 
  mutate(sig=case_when(p_value >= 0.05 ~ "NS",
                       TRUE ~ "*")) %>% 
  filter(sig!="NS") %>% 
  mutate(p_value_mod=case_when(p_value < 0.0001 ~ round(p_value,digits=5),
                               p_value < 0.001 ~ round(p_value,digits=4),
                               p_value < 0.01 ~ round(p_value,digits=3),
                               TRUE ~ round(p_value,digits=2)))

sub_stats <- sub_stats %>% 
  #mutate(y.position=seq.int(0.85,1.05,length.out = nrow(sub_stats))) %>% 
  mutate(y.position=c(1,1,1,1,1,
                      0.9,
                      0.95,0.95,0.95,0.95,
                      0.85,
                      1.05,1.05,1.05,1.05,1.05,1.05)) %>% 
  #mutate(y.position=c(0.85,0.95,0.9,.95)) %>%
  rename(group1=level1,group2=level2)

ggplot(sub_df,aes(x=level,y=AUC)) + 
  geom_hline(yintercept = 0.5,color="grey",lty="dashed") +
  #geom_boxplot(aes(fill=method)) +
  geom_jitter(aes(fill=method),height = 0, width = 0.2,alpha=0.8,shape=21) +
  stat_summary(fun.data = median_hilow,fun.args=0.5, geom="crossbar", width=0.5, color="black") +
  scale_fill_manual(values = pal2[m]) + 
  scale_y_continuous(limits=c(0.4,1.05),breaks = seq(0.4,1.05,0.1)) +
  theme_bw() + xlab("") + ylab("AUROC") +
  theme(legend.position ="none",
        #axis.text.x = element_text(angle = 45,hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.text = element_text(size=14)) +
  #ggtitle(m) +
  #stat_summary(fun = mean, color="white",geom = "point", shape = 18, size = 3,show.legend = FALSE) +
  #stat_pvalue_manual(sub_stats,label="p_value_mod") 
  stat_pvalue_manual(sub_stats,label="sig")
```

@@

::: {.row}
::: {.col-md-6}
```{r glmnet_table,echo=FALSE}
t_table %>%
  select(level,glmnet) %>% 
  mutate(glmnet=round(glmnet,digits = 3)) %>% 
  rename("Median AUC"=glmnet,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
:::

::: {.col-md-6}
```{r glmnet_plot,echo=F,message=F}
## glmnet Model Performance
m <- "Logistic Regression"

@@

```
:::
:::

@@

::: {.row}
::: {.col-md-6}
```{r rpart2_table,echo=FALSE}
t_table %>%
  select(level,rpart2) %>% 
  mutate(rpart2=round(rpart2,digits = 3)) %>% 
  rename("Median AUC"=rpart2,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
:::

::: {.col-md-6}
```{r rpart2_plot,echo=F,message=F}
## rpart2 Model Performance
m <- "Decision Tree"

@@

::: {.row}
::: {.col-md-6}
```{r svmRadial_table,echo=FALSE}
t_table %>%
  select(level,svmRadial) %>% 
  mutate(svmRadial=round(svmRadial,digits = 3)) %>% 
  rename("Median AUC"=svmRadial,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
:::

::: {.col-md-6}
```{r svmRadial_plot,echo=F,message=F}
## svmRadial Model Performance
m <- "SVM Radial"

@@

::: {.row}
::: {.col-md-6}
```{r xgbTree_table,echo=FALSE}
t_table %>%
  select(level,xgbTree) %>% 
  mutate(xgbTree=round(xgbTree,digits = 3)) %>% 
  rename("Median AUC"=xgbTree,"Level"=level) %>% 
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position="center")
```
:::

::: {.col-md-6}
```{r xgbTree_plot,echo=F,message=F}
## XGBoost Model Performance
m <- "XGBoost"

@@

```
:::
:::

## Feature Importance {#feature-importance}

The mikropml packages includes an option for finding [feature importance](http://www.schlosslab.org/mikropml/reference/get_feature_importance.html) using a permutation method. This significantly increases the time to run the models. The function permutes each model feature (e.g. genus) and recalculates the AUC with that feature permuted. The outputs are the permuted AUC value and the difference between this AUC value and the actual AUC value. Larger positive differences can be interpreted as being more important since the AUC value is smaller when this feature is permuted.

@@

### Top 10 Important features

```{r importance_rank, message=F,echo=F}

@@

top_10_tax <- ranked_tax %>% 
  filter(rank <= 10) %>% 
  mutate(Taxonomy=case_when(level=="otu" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            level=="asv" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            TRUE ~ Taxonomy)) %>% 
  mutate(Taxonomy=sapply(strsplit(Taxonomy,"\\("), `[`, 1)) %>% 
  mutate(Taxonomy=paste0(Taxonomy," (",OTU,")"))  %>% 
  select(level,Taxonomy,rank) %>% 
  pivot_wider(names_from = level,values_from = Taxonomy) 
colnames(top_10_tax) <- c("Rank","Phylum","Class","Order","Family","Genus","OTU","ASV")

ranked_tax %>% 
  filter(rank <= 10) %>% 
  mutate(Taxonomy=case_when(level=="otu" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            level=="asv" ~ sapply(strsplit(Taxonomy,"();"), `[`, 6),
                            TRUE ~ Taxonomy)) %>% 
  mutate(Taxonomy=sapply(strsplit(Taxonomy,"\\("), `[`, 1)) %>% 
  mutate(Taxonomy=paste0(Taxonomy," (",OTU,")"))

kable(top_10_tax,format="html",booktabs=T,align="clllllll",
      col.names = c("Rank",levels_names)) %>% 
   kable_material(c("hover","condensed"))

### EXAMPLE plot
library(glue)
library(ggtext)

data <- tibble(bactname = c("Staphylococcaceae", "Moraxella", "Streptococcus", "Acinetobacter"),
               OTUname = c("OTU 1", "OTU 2", "OTU 3", "OTU 4"),
               value = c(-0.5, 0.5, 2, 3))

data %>% mutate(
  name = glue("<i>{bactname}</i> ({OTUname})"),
  name = fct_reorder(name, value))  %>%
  ggplot(aes(value, name)) + 
    geom_col(alpha = 0.5) + 
    scale_fill_identity() +
    theme(
      axis.text.y = element_markdown(),
      plot.caption = element_markdown(lineheight = 1.2))

## Prevalence {#prevalence}

```{r prevalence,echo=F,message=F}

prevalence <- read_csv("../analysis/prevalence_by_level.csv")

prevalence$level <- factor(prevalence$level,levels=levels)

tax_count <- prevalence %>% 
  filter(kept==1) %>% 
  group_by(level) %>% 
  mutate(numOTU = n_distinct(name)) %>% 
  select(level,numOTU) %>% 
  distinct()

pct_bins <- NULL
level_nOTU <- NULL
for(l in levels){
  sub_data <- prevalence %>% 
    filter(level==l,kept==1)
  hist <- hist(sub_data$prevalence,breaks = seq(0,1,0.25),plot=FALSE)

@@

level_nOTU <- level_nOTU %>% 
  mutate(level_name = levels_names) 
labs <- paste0(level_nOTU$level_name," (",level_nOTU$nOTU,")")
pct_bins %>%
  mutate(level=factor(level,levels=levels,labels=labs)) %>% 
  ggplot(aes(x=prev,y=pct,fill=level)) +
    geom_bar(stat = "identity") +
    facet_wrap(~level) +
    scale_y_continuous(expand=c(0,0),limits=c(0,0.8),labels = scales::percent_format()) +
    scale_x_continuous(expand=c(0,0),breaks = c(0,.25,0.50,.75,1)) +
    xlab("Prevalence in Samples") +
    ylab("Percent of Taxa") +
    theme_bw() +
    theme(legend.position = "none",
          axis.text = element_text(size=12),
          axis.title = element_text(size=14)) +
  geom_text(aes(label = paste0((round(pct,digits=2)*100),"%")),nudge_y=0.05)
ggsave("./figures/bin_prevalence.png")
```
